---
title: "hw04"
author: "Wynes"
date: "October 9, 2017"
output: github_document
---
Part I - Importing and tidying a database

For this exercise I want to import a database and then attach it to gapminder. The database is of carbon dioxide per capita emissions from: EDGARv4.2, European Commission, Joint Research Centre (JRC)/PBL Netherlands Environmental Assessment Agency. Emission Database for Global Atmospheric Research (EDGAR), release version 4.2. http://edgar.jrc.ec.europe.eu, 2011

So I'm using the read_csv function from the "readr" package, which is already part of the tidyverse. This will produce a tibble from the csv.

```{r}

library("tidyverse")

co2percap <- read_csv("C:/Users/AsusW10/Documents/STAT545/co2pcap.csv")
View(co2percap)
```


Now because the years are listed as columns and the countries are listed as rows I need to rearrange the data.

```{r}
co2percap %>% gather(year,co2pcap,-c(Country)) %>% 
  arrange(Country)
  knitr::kable()
```
So I've successfully converted a wide data frame into a long data frame and then used the knitr::kable function to display this in a pleasing way.

Let's graph the data to see that it looks reasonable and that nothing has gone amiss.
```{r}
ghg <- co2percap %>% gather(year,co2pcap,-c(Country)) %>% 
  arrange(Country) 
ghg %>% 
  filter(!is.na(co2pcap)) %>% 
  ggplot(aes(x=year, y=co2pcap))+
  geom_point(alpha=0.2)
```
There are a lot of NA values, and I tried removing them with na.rm=T, and na.omit but nothing seems to be working. The data seems fine when examining it in a table, but smell testing it with ggplot isn't going to work. I'm going to look at Canadian values for the years where data is available.

```{r}
ghg %>% 
  filter(Country=="Canada",year<2014, year>1970) %>% 
  ggplot(aes(x=year, y=co2pcap)) +
  geom_point()+ 
  theme(axis.text.x=element_text(angle=90,hjust=1))
```
These values range within a reasonable space for Canadian per capita emissions.

##Part II - Joining with gapminder using dplyr

When I first tried to join these datasets I realized that the column name for country was capitalized in the ghg dataset but not in gapminder, so I used the rename function so that dplyr would recognize the columns as being the same. After this 
```{r}
ghg2 <- rename(ghg,country = Country)
View(ghg2)

ghg4 <- ghg2[complete.cases(ghg2), ]

```


So I tried to use this code

left_join(gapminder, ghg4, by = c("country","year"))

to join the dataframes, but have come up with an error because the year column in the ghg dataframe is not integers. But I found a solution that seems to work on stackoverflow (https://stackoverflow.com/questions/43126096/change-select-columns-from-character-to-integers)
```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
left_join(gapminder, ghg4, by = c("country","year"))
```
Because gapminder only has data for ever five years, four out of every five rows are blank for the gapminder data. I can try a left join using the co2 data as my left dataframe

```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
right_join(gapminder, ghg4, by = c("country","year"))
```
This time I noticed that the co2 data does not go back as far, so there are NA values for 1952, 1957, 1962, and 1967 for the new co2pcap column in this dataframe.

```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
inner_join(gapminder, ghg4, by = c("country","year"))
```
This dataframe only contains rows where I have full information from both datasets. Depending on one's purposes this might be the easiest data to explore. I'm going to perform a quick test with it:

```{r}
 ghg_gap <- ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
inner_join(gapminder, ghg4, by = c("country","year"))

ghg_gap %>% 
  ggplot(aes(x=lifeExp,y=co2pcap), na.rm=TRUE)+
  geom_point(alpha=0.2)

```
So something seems to be wrong with the data but I"m not sure what. I've seen this before with NA values but I don't have any here. I'll try some other joins and see if they experience the same problems.

```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
full_join(gapminder, ghg4, by = c("country","year"))
```
I tried semi_join and found that it really returned back very little information. full_join, as seen above, retains all of the data from both dataframes, with the downside that there are now a lot of NA values.

#Reflection
So I know I accomplished a lot less than I normally would because I chose a dataframe that wasn't nicely cleaned up already. It took me some time to find out how to convert a csv into a tibble (among many resources, I used this page: http://readr.tidyverse.org/reference/read_delim.html). 

Additionally, it was very difficult to join the two dataframes. The dataframe on CO2 per capita emissions by different countries has NA values which seems to cause errors when joining them. I managed to use complete.cases to create a version of the database without NA values (in actual research this would not be ideal but I am just trying to learn how to join dataframes). Even then I was still getting error messages for having incompatible types of columns which took more troubleshooting.