---
title: "hw04"
author: "Wynes"
date: "October 9, 2017"
output: github_document
---
#Part I - Importing and tidying a database

For this exercise I want to import a database and then attach it to gapminder. The database is of carbon dioxide per capita emissions from: EDGARv4.2, European Commission, Joint Research Centre (JRC)/PBL Netherlands Environmental Assessment Agency. Emission Database for Global Atmospheric Research (EDGAR), release version 4.2. http://edgar.jrc.ec.europe.eu, 2011

So I'm using the read_csv function from the "readr" package, which is already part of the tidyverse. This will produce a tibble from the csv.

```{r}

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))

co2percap <- read_csv("C:/Users/AsusW10/Documents/STAT545/co2pcap.csv")
head(co2percap)
```



Now because the years are listed as columns and the countries are listed as rows I need to rearrange the data.

```{r}
co2percap %>% gather(year,co2pcap,-c(Country)) %>% 
  arrange(Country)
```



So I've successfully converted a wide data frame into a long data frame.

Let's graph the data to see that it looks reasonable and that nothing has gone amiss.



```{r}
ghg <- co2percap %>% gather(year,co2pcap,-c(Country)) %>% 
  arrange(Country) 
ghg %>% 
  filter(!is.na(co2pcap)) %>% 
  ggplot(aes(x=year, y=co2pcap))+
  geom_point(alpha=0.2)
```



There are a lot of NA values, and I tried removing them with na.rm=T, and na.omit but nothing seems to be working. The data seems fine when examining it in a table, but smell testing it with ggplot isn't going to work. I'm going to look at Canadian values for the years where data is available.


```{r}
ghg %>% 
  filter(Country=="Canada",year<2014, year>1970) %>% 
  ggplot(aes(x=year, y=co2pcap)) +
  geom_point()+ 
  theme(axis.text.x=element_text(angle=90,hjust=1))
```



These values range within a reasonable space for Canadian per capita emissions.

##Part II - Joining with gapminder using dplyr

When I first tried to join these datasets I realized that the column name for country was capitalized in the ghg dataset but not in gapminder, so I used the rename function so that dplyr would recognize the columns as being the same.


```{r}
ghg2 <- rename(ghg,country = Country)
View(ghg2)

ghg4 <- ghg2[complete.cases(ghg2), ]

```



Next, I tried to use this code

left_join(gapminder, ghg4, by = c("country","year"))

to join the dataframes, but have come up with an error because the year column in the ghg dataframe is not integers. But I found a solution that seems to work on stackoverflow (https://stackoverflow.com/questions/43126096/change-select-columns-from-character-to-integers)


```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
left_join(gapminder, ghg4, by = c("country","year"))
```



Because gapminder only has data for every five years, four out of every five rows are blank for the gapminder data. I can try a right join instead.


```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
right_join(gapminder, ghg4, by = c("country","year"))
```



This time I noticed that the co2 data does not go back as far, so there are NA values for 1952, 1957, 1962, and 1967 for the new co2pcap column in this dataframe. Next I will try inner_join


```{r}
ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
    #The second column is year, the one we want to mutate into a numeric value, which is why c(2) is specified. I could specify more columns by adding other numbers.
  inner_join(gapminder, ghg4, by = c("country","year"))

```



This dataframe only contains rows where I have full information from both datasets. Depending on one's purposes this might be the easiest data to explore. I'm going to perform a quick test with it:


```{r}
#First the same code as before, but placed into an object that we can graph 
ghg_gap <- ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
inner_join(gapminder, ghg4, by = c("country","year"))
View(ghg_gap)

#Now graph that object ("ghg_gap")
ghg_gap %>% 
  ggplot(aes(x=lifeExp,y=co2pcap))+
  geom_point(alpha=0.2)


```



So something seems to be wrong with the data but I"m not sure what. I've seen this type of graphical error occur before when NA values were present in the data but I don't have any here (I checked with the is.na function). Even working with a TA I was unable to solve this. 

Below I try another join.


```{r}
 ghg4 %>% 
  mutate_at(c(2), as.numeric) %>%
full_join(gapminder, ghg4, by = c("country","year"))
```



I tried semi_join and found that it really returned back very little information. full_join, as seen above, retains all of the data from both dataframes, with the downside that there are now a lot of NA values.

#Reflection
Using a dataframe that wasn't already cleaned up and ready for R took a lot of time. It was difficult to find a dataset that appropriately matched gapminder and then to find out how to convert a csv into a tibble (among many resources, I used this page: http://readr.tidyverse.org/reference/read_delim.html).

Additionally, it was very difficult to join the two dataframes. The dataframe on CO2 per capita emissions by different countries has NA values which seems to cause errors when joining them. I managed to use complete.cases to create a version of the database without NA values (in actual research this would not be ideal but I am just trying to learn how to join dataframes). Even then I was still getting error messages for having incompatible types of columns which took more troubleshooting. But I'm excited because these are problems I've faced in the past and now I have at least some answers.