---
title: "carbonfly vignette"
author: "Seth Wynes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction to carbonfly

Determing the emissions associated with air travel between two locations can be difficult. Many websites are available to do this for you, but it can be hard to pass your data through the websites, and they are not always transparent about their methodology.

This package:

- Allows you to calculate greater circle distance between two coordinates with great accuracy
- Allows you to calculate air travel emissions if you know absolute distance
- Allows you to calculate air travel emissions between two airports using the IATA 3-letter code.

## Greater Circle Distance

The gcd() function allows you to calculate greater circle distance between two points on the globe, as measured in kilometers. Because the globe is spherical, when travelling across it you cannot use a straight line to calculate distance between two points. Additionally, the globe is not a perfect sphere, so complicated trigonometry is required to produce accurate calculations. The gcd() function takes all of this into account. For longer methods this is more accurate (though it breaks down for antipodal points). Code adapted from: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/

The function takes the general form gcd(long1,lat1,long2,lat2) where
- long1 and lat1 are the longitude and latitude coordinates of the starting point
- long2 and lat2 are the longitude and latitude coordinates of the end point

**NOTE** In this formula longitude comes first, though latitude is usually listed first by convention

**EXAMPLE**
Calculation for the distance between the Eiffel tower in Paris (48.8584 degrees North, 2.2945 degrees East) and the Great Pyramid of Giza (29.9792 degrees North, 31.1342 degrees East). Note that latitudes south of the equator are negative and that longitudes west of the Prime Meridian are negative. 

```{r}
library(carbonfly)
gcd(48.8584,2.2945,29.9792,31.1342)
gcd(2.2945,48.8584,31.1342,29.9792)
```



**EXAMPLE**
Fly from Toronto (43.6777째 N, 79.6248째 W) to Johannesburg airport (JNB) (26.1367째 S, 28.2411째 E)

```{r}
gcd(-79.6248,43.6777,28.2411,-26.1367)
```



## Calculating emissions when you know distance

The CO2calclulator() function can be used when you know the distance between the two points you are travelling. The emissions factors that inform this calculation take into account the fact that short trips use more fuel per kilometer than long trips because takeoffs burn more fuel than cruising. The emissions factors also account for average aircraft occupancy so the figure provided represents the total of greenhouse gases burned per passenger.

Emissions factors:
flights <463km 0.27867kgCo2e per passenger per kilometer

Distance | Emissions Factors
------------- | -------------
<463km | 0.27867kgCo2e pkm
463-3700km | 0.16844kgCO2e pkm
>3700km | 0.19162kgCO2e pkm

**EXAMPLES**
Find emissions for a 100km flight

```{r}
co2calculator(100)
```



## Calculating emissions from two airport codes

The carboncalculator() function takes two inputs, an origin airport and a destination airport and returns the total emissions per passenger created by taking the flight as measured in kgCO2e.

**EXAMPLE**
```{r}
library(dplyr)
library(magrittr)
carboncalculator("YYZ","YVR")
```

