---
title: "hw05"
author: "Wynes"
date: "October 14, 2017"
output: html_document
---
In this assignment I'm practicing working with factors and inputting and outputting the data. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(forcats))
```



*Dropping factors


Drop Oceania from the list of continents
```{r}
#Filter out Oceania
levels(gapminder$continent)
noo_continents <- c("Asia", "Africa", "Europe", "Americas")
noo_gap <- gapminder %>%
  filter(continent %in% noo_continents)

#Now remove the unwanted Oceania level from the new df h_gap
noo_gap_drop <- noo_gap %>% 
  droplevels()
```



So I've removed the Oceania continent in the new object, noo_gap, and then in the object noo_gap_drop I've gone further and dropped the levels associated with it. We can inspect both of the objects to contrast their properties.

```{r}
nlevels(noo_gap$continent)
nlevels(noo_gap_drop$continent)

noo_gap$continent %>% 
levels() 

noo_gap_drop$continent %>% 
  levels()
  
fct_count(noo_gap$continent)
fct_count(noo_gap_drop$continent)

```



From the levels function we see that noo_gap has one more continent listed as a factor than noo_gap_drop and two more countries.

From the fct_count argument we can see that noo_gap has the same number of rows for each continent as noo_gap_drop except that Oceania is missing completely in noo_gap_drop whereas in noo_gap it is listed as having zero rows.

*Rearranging Factors
```{r}
noo_gap_drop$continent %>% 
  fct_infreq() %>%
  levels() %>% head()
```



Here I've rearranged so that the factor with the largest number of rows (see the fct_count result above) comes first.

This isn't terribly interesting so let's rearrange countries based on GDP per capita instead.


```{r}
fct_reorder(noo_gap_drop$country, noo_gap_drop$gdpPercap) %>% 
  levels() %>% head()

#The default uses median GDP per capita. Here I update to using mean:
fct_reorder(noo_gap_drop$country, noo_gap_drop$gdpPercap, mean) %>% 
  levels() %>% head()

#And now use reordering inside ggplot
noo_gap_drop2 <- noo_gap_drop %>% filter(continent == "Americas", year =="2007")
ggplot(noo_gap_drop2, aes(y = gdpPercap, x = fct_reorder(country, gdpPercap))) +
  geom_col() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(x="Country" , y= "GDP per capita")

```



That worked well, let's try and replicate that with the arrange function:
```{r}
  noo_gap_drop2 %>% 
  arrange(gdpPercap) %>% 
ggplot(aes(y = gdpPercap, x = country)) +
  geom_col() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(x="Country" , y= "GDP per capita")
```



I conclude that arrange does not carry over into ggplot with piping. I'm going to add the fct_reorder function back in and see if it is still compatible with the arrange function, but give arrange an opposing command (to arrange by life expectancy instead of GDP per capita).

```{r}
  noo_gap_drop2 %>% 
  arrange(lifeExp) %>% 
ggplot(aes(y = gdpPercap, x = fct_reorder(country, gdpPercap))) +
  geom_col() +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(x="Country" , y= "GDP per capita")
```



And the figure was clearly ordered by GDP per capita with no regards for the arrange function.

Here I check and see that arrange is still working, and find that it does order the countries correctly by GDP per capita in a table.

```{r}
 noo_gap_drop2 %>% 
  arrange(gdpPercap) %>% 
  filter(continent=="Americas") %>% 
  glimpse() %>% 
  knitr::kable()
```



*File I/O
Now that the data has had Oceania removed, levels dropped, filtered to just the Americas and arranged by GDP per capita I'm going to write it to file and then import it back to see what changes have occurred.

```{r}
America_GDP <- noo_gap_drop2 %>% 
  arrange(gdpPercap) %>% 
  filter(continent=="Americas")

head(America_GDP) %>% 
  knitr::kable()
```



Writing to file:
```{r}
write_csv(America_GDP, "America_GDP.csv")
```



At this point I checked and found a csv file in my working directory on my laptop. Now to read the file back in.

```{r}
America_GDP2 <- read_csv("America_GDP.csv")

glimpse(America_GDP2) %>% 
  knitr::kable
```



Recalling the original object (see below) before the input output process, we can see that the order of the countries has been preserved, but in the original object country and continents were factors and in the imported object they are both coded as characters.
```{r}
glimpse(America_GDP) %>% 
  knitr::kable
```

