---
title: "hw05"
author: "Wynes"
date: "October 14, 2017"
output: html_document
---
In this assignment I'm practicing working with factors and inputting and outputting the data. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(forcats))
```



*Dropping factors


Drop Oceania from the list of continents
```{r}
#Filter out Oceania
list(gapminder$continent)
noo_continents <- c("Asia", "Africa", "Europe", "Americas")
noo_gap <- gapminder %>%
  filter(continent %in% h_continents)

#Now remove the unwanted Oceania level from the new df h_gap
noo_gap_drop <- noo_gap %>% 
  droplevels()
```



So I've removed the Oceania continent in the new object, noo_gap, and then in the object noo_gap_drop I've gone further and dropped the levels associated with it. We can inspect both of the objects to contrast their properties.

```{r}
nlevels(noo_gap$continent)
nlevels(noo_gap_drop$continent)

noo_gap$continent %>% 
levels() 

noo_gap_drop$continent %>% 
  levels()
  
fct_count(noo_gap$continent)
fct_count(noo_gap_drop$continent)

```



From the levels function we see that noo_gap has one more continent listed as a factor than noo_gap_drop and two more countries.

From the fct_count argument we can see that noo_gap has the same number of rows for each continent as noo_gap_drop except that Oceania is missing completely in noo_gap_drop whereas in noo_gap it is listed as having zero rows.

*Rearranging Factors
```{r}
noo_gap_drop$continent %>% 
  fct_infreq() %>%
  levels() %>% head()
```



Here I've rearranged so that the factor with the largest number of rows (see the fct_count result above) comes first.

This isn't terribly interesting so let's rearrange countries based on GDP per capita instead.


```{r}
fct_reorder(noo_gap_drop$country, noo_gap_drop$gdpPercap) %>% 
  levels() %>% head()

#The default uses median GDP per capita. Here I update to using mean:
fct_reorder(noo_gap_drop$country, noo_gap_drop$lifeExp, mean) %>% 
  levels() %>% head()

#And now 

```


