---
title: "hw06"
output: html_document
---

```{r setup, cache = F}
knitr::opts_chunk$set(error = TRUE)
#This alters the knit process so that errors can be displayed, which I need for this session where I intentionally introduce errors to test my functions
```

### Practicing functions with a CARBON CALCULATOR

In this homework I'm going to start by trying to make my own function that acts as a carbon calculator (translating two airport codes into a distance and then an amount of greenhouse gases).

Step 1 is to find code that converts latitude and longitude into greater circle distance. This code is long, but is very precise because it accounts for the fact that the earth is not a perfect sphere, which most available calculators neglect. I found it here: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/
```{r}
deg2rad <- function(deg) return(deg*pi/180) #Change degrees into radians
#CHECK
deg2rad(180)

gcd.vif <- function(long1, lat1, long2, lat2) {

# WGS-84 ellipsoid parameters
a <- 6378137         # length of major axis of the ellipsoid (radius at equator)
b <- 6356752.314245  # ength of minor axis of the ellipsoid (radius at the poles)
f <- 1/298.257223563 # flattening of the ellipsoid

L <- long2-long1 # difference in longitude
U1 <- atan((1-f) * tan(lat1)) # reduced latitude
U2 <- atan((1-f) * tan(lat2)) # reduced latitude
sinU1 <- sin(U1)
cosU1 <- cos(U1)
sinU2 <- sin(U2)
cosU2 <- cos(U2)

cosSqAlpha <- NULL
sinSigma <- NULL
cosSigma <- NULL
cos2SigmaM <- NULL
sigma <- NULL

lambda <- L
lambdaP <- 0
iterLimit <- 100
while (abs(lambda-lambdaP) > 1e-12 & iterLimit>0) {
sinLambda <- sin(lambda)
cosLambda <- cos(lambda)
sinSigma <- sqrt( (cosU2*sinLambda) * (cosU2*sinLambda) +
(cosU1*sinU2-sinU1*cosU2*cosLambda) * (cosU1*sinU2-sinU1*cosU2*cosLambda) )
if (sinSigma==0) return(0)  # Co-incident points
cosSigma <- sinU1*sinU2 + cosU1*cosU2*cosLambda
sigma <- atan2(sinSigma, cosSigma)
sinAlpha <- cosU1 * cosU2 * sinLambda / sinSigma
cosSqAlpha <- 1 - sinAlpha*sinAlpha
cos2SigmaM <- cosSigma - 2*sinU1*sinU2/cosSqAlpha
if (is.na(cos2SigmaM)) cos2SigmaM <- 0  # Equatorial line: cosSqAlpha=0
C <- f/16*cosSqAlpha*(4+f*(4-3*cosSqAlpha))
lambdaP <- lambda
lambda <- L + (1-C) * f * sinAlpha *
(sigma + C*sinSigma*(cos2SigmaM+C*cosSigma*(-1+2*cos2SigmaM*cos2SigmaM)))
iterLimit <- iterLimit - 1
}
if (iterLimit==0) return(NA)  # formula failed to converge
uSq <- cosSqAlpha * (a*a - b*b) / (b*b)
A <- 1 + uSq/16384*(4096+uSq*(-768+uSq*(320-175*uSq)))
B <- uSq/1024 * (256+uSq*(-128+uSq*(74-47*uSq)))
deltaSigma = B*sinSigma*(cos2SigmaM+B/4*(cosSigma*(-1+2*cos2SigmaM^2) -
B/6*cos2SigmaM*(-3+4*sinSigma^2)*(-3+4*cos2SigmaM^2)))
s <- b*A*(sigma-deltaSigma) / 1000

return(s) # Distance in km
}
```



Now I run the code with example coordinates (translated into radians) for the flight between Pearson and Gatwick airport:

```{r}
#Template for GCD
#gcd.vif(long1,lat1,long2,lat2)

#Actual calculation
gcd.vif(-1.389815387,0.762311349,-0.003320978,0.892701983)
```



Now I want a function that will take that distance, and multiply it by a certain amount of emissions according to some if statements. So if the distance is less than 463km it should apply one emissions factor (0.27867kgCO2/km), if it is longer than 463 but shorter than 3700 it should apply a second emissions factor (0.16844kgCO2/km), and if it is longer than or equal to 3700 it should multiply the distance by a third emissions factor(0.19162kgCO2/km). Emissions factors from: (https://www.gov.uk/government/publications/greenhouse-gas-reporting-conversion-factors-2016)
R operators from: https://www.statmethods.net/management/operators.html


```{r}
co2calculator <- function(distance) {
  co2e<- if(distance<=463){
    distance*0.27867
    }   else if (distance>463 && distance<3700) {
    distance*0.16844
  } else if (distance>=3700) {
    distance*0.19162
  } 
  return(co2e)
  }

co2calculator(1) # Check that for a value under 463 0.27867 is the emissions factor
co2calculator(1000) #Check that for a medium range value 0.16844 is the factor
co2calculator(10000) #Check that for a long range value 0.19162 is the factor
```



Now if I nest the function for changing longitude and latitude into radians, inside of the function for getting distance from radians, inside of my carbon calculator, I can get the results by inputting only the longitude and latitude coordinates. Let's call this function carboncalculator and test it.

```{r}
  #note that I have switched  the order of lat and long in this formula because the #function I copied from the internet went against geography conventions
carboncalculator <- function(lat1deg,long1deg,lat2deg,long2deg) {
  co2calculator(
    gcd.vif(
      deg2rad(long1deg),
      deg2rad(lat1deg),
      deg2rad(long2deg),
      deg2rad(lat2deg)))
}

carboncalculator(50,100,50,100)
#This returns an error which is good because the length of the trip is zero

carboncalculator(49.3,-123.1,50,-119.4)
#This is the trip from Vancouver to Kelowna, and the answer provided is reasonable (77kg of CO2)

carboncalculator(51.2,0.2,22.3,113.9)
#This is the trip from London (Gatwick Airport) to Hong Kong, and the answer provided is reasonable (1844kg of CO2)

carboncalculator(TRUE,TRUE,FALSE,FALSE)
#Amazingly, this gives us an answer, which is a problem. I now need to add some safeguards.

```



Adding stop if statements to my carbon calculator function:

```{r}
carboncalculator2 <- function(lat1deg,long1deg,lat2deg,long2deg) {
    if(!is.numeric(lat1deg)){
      stop('this function only works for numeric input\n',
           'You have provided an object of class:',class(lat1deg)[1])
    }
    co2calculator(
    gcd.vif(
      deg2rad(long1deg),
      deg2rad(lat1deg),
      deg2rad(long2deg),
      deg2rad(lat2deg)))
}

carboncalculator2(1,2,3,4)
#Check that it still works the same way for regular inputs

carboncalculator2(TRUE,TRUE,FALSE,FALSE)
#This gives the appropriate error message, but now I need to add that error message for all longitude and latitude inputs:

carboncalculator3 <- function(lat1deg,long1deg,lat2deg,long2deg) {
    if(!is.numeric(lat1deg)){
      stop('this function only works for numeric input\n',
           'For latitude 1, you have provided an object of class:',class(lat1deg)[1])
    }
   if(!is.numeric(long1deg)){
      stop('this function only works for numeric input\n',
           'For longitude 1, you have provided an object of class:',class(long1deg)[1])
    }
   if(!is.numeric(lat2deg)){
      stop('this function only works for numeric input\n',
           'For latitude 2, you have provided an object of class:',class(lat2deg)[1])
    }
   if(!is.numeric(long2deg)){
      stop('this function only works for numeric input\n',
           'For longitude 2, you have provided an object of class:',class(long2deg)[1])
    }
      co2calculator(
    gcd.vif(
      deg2rad(long1deg),
      deg2rad(lat1deg),
      deg2rad(long2deg),
      deg2rad(lat2deg)))
}

carboncalculator3(1,2,3,4)
#Check to see that it still works on regular inputs

carboncalculator3(1,TRUE,3,4)


```



I now have an error message that tells you EXACTLY which input is of the wrong class and which class it should be! 

The next step for my carbon calculator would be to have it work just by inputting airport codes instead of longitude and latitude. I will have to leave that for my package because the first half took so long.


## Working with the Singer Data
```{r}
library(singer)
library(tidyverse)
library(purrr)
library(repurrrsive)
library(ggmap)
?revgeocode

```

```{r}
#Practice using revgeocode with a simple example of Vancouver
head(singer_locations)
revgeocode(location=c(-79.3832,43.6532), output=c("more"))

```



This functions, and provides a lot of detail. Next I'll need to clean up the singer_locations database so it's ready for this function. Because they're not useful, and I don't want to waste the 2500 query limit per day for this function, I'm going to start by omiting all of the NA values. Furthermore, I'm going to drop all of the songs not from 2007 so that there is a more manageable number of songs to practice getting this code right with.

Sidenote: While trying out how to solve this problem I read a great blog on how R uses vectors that I really reccommend: http://www.noamross.net/blog/2014/4/16/vectorization-in-r--why.html


```{r}
  levels(singer_locations$year)

#Create an object with the years I want
late_singer <- c("2006","2007")

late_singer_locations <- singer_locations %>%
  filter(year %in% late_singer)

head(late_singer_locations)

#Now remove the unwanted levels associated with the omitted data
late_sing_loc <- late_singer_locations %>% 
  droplevels()

View(late_sing_loc)

late_sing_loc %>% 
  na.omit() %>% 
  revgeocode(location=c(late_sing_loc$longitude,late_sing_loc$latitude), output=c("more")) #This doesn't work because revgeocode wants a vector and not a column in a dataframe, I guess. 


late_sing_loc %>% 
  na.omit() %>% 
sapply(11:12, revgeocode(location=c(longitude,latitude), output=c("more")), simplify = TRUE)
#Here I use sapply and specify the eleventh and twelfth columns with no success.

sapply(late_sing_loc[11:12], revgeocode(location=c(late_sing_loc$longitude,late_sing_loc$latitude), output=c("more")), simplify = TRUE)
#And above I tried the same, but without piping. No luck.


#To make my code simpler I create two new vectors which are the latitude and longitude columns without the NA values. I now know they are the same length and they are doubles, so less errors are now likely to arise.
longi <-late_sing_loc$longitude %>% na.omit()
lati <- late_sing_loc$latitude %>% na.omit()
typeof(lati) #New vectors are doubles

map2(longi,lati, revgeocode(location=c(longi,lati), output=c("more")), simplify = TRUE)

#Below I try instead to make a new function that is revgeocode but simplified:
revgeocode2 <- function(longi,lati) {
  revgeocode(location=c(longi,lati), ouput=c("address"))
}


mapply(revgeocode2(longi,lati),longi,lati)
#It also fails to work, even with mapply

mapply(revgeocode2,longi,lati)
#Similar failure
```



So at this point I will settle for having a function with two arguments that works in *apply with my two vectors. Below I make that arbitrary function and try to use it with the apply family of functions.

```{r}
latlong <- function(longi,lati) {
  3*longi*lati
}


#Testing that function works
latlong(1,1)
latlong(2,4)


#Using apply with the new function.
check1latlong <- mapply(latlong,longi,lati)
#And this successfully applied the function with two arguments to every entry in the two vectors. So the problem must be with the function revgeocode.

check2latlong <- map2(longi,lati,latlong)
#This code also worked to generate the same result.

head(check1latlong)
head(check2latlong)

#These solutions came from a stackoverflow problem, but were not successful in my specific case of wanting to use revgeocode: https://stackoverflow.com/questions/14427253/passing-several-arguments-to-fun-of-lapply-and-others-apply
```


## Reflection on the process

This was the hardest assignment for me so far. I learned a lot about functions, but it took so much code to implement them correctly that I couldn't quite make my target of a completed carbon calculator. But I was glad to create my own, useful function that includes warning errors.

In the second half of the assignment I am stuck with some glitch that won't allow me to use the revgeocode function in conjunction with the apply family of functions and despite talking with a TA about it I was unable to solve it. I am glad for the experience with the apply functions, however, as I now know that I can use them to apply my carbon calculator function to a dataframe.
